Type Refinements for the Working Class

### 2015-12-31 paper.tex
Define symbol contexts, valences, arities, operators

See~\cite{sterling-morrison:2015} for the development of abstract binding trees
with symbols. \todo[inline]{give a brief description of the framework, and present its
rules.}

>>>>

We will briefly reproduce the syntax of abts here;
see~\cite{sterling-morrison:2015} for a full development of the syntax and
semantics of multi-sorted nominal abstract binding trees, however. The point
of nominal abts is that they extend the standard treatments of second order
universal algebra~\cite{fiore-hur:2010} with a notion of \emph{symbolic
parameter} and name abstraction; this is absolutely essential for correctly
treating the syntax of assignables, open exceptions, fluid bindings, and
processes, among other things. See Harper's \emph{Practical Foundations for
Programming Languages}~$\cite{harper:2016}$ for numerous examples of the use of
symbolic parameters.

Fix a set of \emph{sorts} $\Sorts$; we'll say $\IsSort{\tau}$ when
$\Member{\tau}{\Sorts}$. Let $\SymSets$ be the category of finite cardinals and
injective maps between them; using the comma construction, we can define the
category $\SCtx$ of \emph{contexts} of symbols, whose objects are finite sets
of symbols $U$ and sort-assignments
\begin{tikzcd}[cramped]
  \IMode{U}\arrow[r, "\IMode{\mathfrak{s}}"] &\IMode{\Sorts}
\end{tikzcd}, and whose morphisms are
sort-preserving renamings; we will write $\Upsilon$ for a symbol context
$(U,\mathfrak{s})$.

We will say that an $\Sorts$-valence is a form
$\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}$, where
$\IMode{\vec{\sigma}\cup\vec{\tau}}\subseteq\IMode{\Sorts}$ and
$\IsSort{\tau}$; valences represent the \emph{binding structure} of an
abstraction of sort $\tau$, where $\vec{\sigma}$ are the sorts of the bound
symbols and $\vec{\tau}$ are the sorts of the bound variables. When $v$ is a
valence, we will say $\IsValence{v}$. An $\Sorts$-arity is of the form
$\MkArity{\vec{v}}{\tau}$, where $\vec{v}$ are $\Sorts$-valences and
$\IsSort{\tau}$; arities classify \emph{operators} (also called ``function
symbols'' in older treatments of algebra). When $a$ is an arity, we will say
$\IsArity{a}$.

An \emph{operator signature} is an arity-indexed family of copresheaves on
$\SCtx$, namely $\HypJ{\Of{\Operators_a}{\Sets^\SCtx}}{\IsArity{a}}$; this is
an intensional generalization of the state of affairs for standard universal
algebra, where operators are arity-indexed families of \emph{sets}. The point
of fibering over symbol contexts is to introduce an open-ended indexing
mechanism for operators; the functorial liftings of $\SCtx$-morphisms
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
  \end{tikzcd}
equip operators $\Member{\vartheta}{\Operators_a(\Upsilon)}$ with a suitable
renaming action $\Member{\Rename{\rho}{\vartheta}}{\Operators_a(\Upsilon')}$.%
%
\footnote{
  In~\cite{sterling-morrison:2015}, the authors defined operator signatures as
  \emph{sheaves} on $\OpCat{(\SCtx)}$ under the atomic topology (alternatively,
  pullback-preserving presheaves), where all non-empty sieves cover; the sheaf
  condition induces the unique existence of \emph{strengthenings} for
  operators. However, because we do not use strengthening in the present
  development, we have omitted this part of the apparatus for the sake of
  simplicity.
}

% \todo[inline]{give a brief description of the framework, and present its rules.}
